Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-09-13 writeup/pnas-word-order-si.tex

We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding the objections by \citet{dunn2011evolved}, we refer to the follow-ups by \citet{levy2011computational}, and \citet{croft2011greenbergian}.}
Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
\citet{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
While most of the other relations also correspond to UD relations, some are not annotated reliably.
We were able formalize eleven out of Dryer's sixteen correlations in UD.
Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.

@@

\paragraph{Evaluating Accuracy of Formalization}
An anonymous reviewer notes that the mapping between Dryer's relations and UD is not perfect, since some of the UD relations subsume other relations.
Here we show that this is not a problem, since the ordering of the various relations subsumed under the UD label typologically strongly agree.

@@

We used the UD treebanks to confirm that different auxiliaires tend to pattern together, and that the most frequent order of the \emph{aux} relation coincides with that of inflected tense-aspect or negation auxiliaries.

We collected, for each UD language, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders, which we operationalized as their more common order in the treebank (auxiliary-head or head-auxiliary).
The dependency occurs in all but two very small treebanks (Telugu and Irish).
In 43 languages, all extracted auxiliaries had the same dominant order, with the possible exception of uninflected particles labeled \emph{aux} (Croatian, German, Polish, Ukrainian).
In three languages (Ancient Greek, Russian), there were other auxiliaries with different dominant order, but these were modal or passive auxiliaries.
Finally, in three languages (Afrikaans, Old Church Slavonic, and Persian), not all tense-aspect auxiliaries showed the same dominant order as the \emph{aux} dependency overall.
For instance, in Persian, the perfect auxiliary \emph{budan} follows the main verb, whereas the future auxiliary \emph{xaastan xaah-} precedes it~\citep[pp. 117, 121]{mace2015persian}.
%		(1) In Afrikaans, the past auxiliary \emph{het} shows dominant aux-head order, whereas the overall most common order of auxiliaries is head-aux (e.g. future auxiliary \emph{sal}).
%		Finally, in Old Church Slavonic, most aux dependencies have the reflexive \emph{sebe} which tends to precede the head, whereas inflected verbal auxiliaries (primarily, forms of \emph{byti}) have opposite order, leading to a seeming violation of the correlation  \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.

@@

\section{Formalizing Comunicative Efficiency}
\subsection{Derivation and Relation to Other Work}

@@

Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms:
\begin{equation}\label{eq:eff-basic}
\text{Informativity} - \lambda \cdot \text{Complexity}
\end{equation}
with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely,hahn2018information, peloquin2019interactions,zaslavsky2019semantic}.

\paragraph{Derivation of our Formalization}
The basis for our precise formalization is the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
%The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms. %; there are some differences in the formalization of complexity.
If $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
\begin{equation}\label{eq:eff-general}
\operatorname{I}[S, R] - \lambda \cdot \operatorname{H}[S]
\end{equation}
where $\operatorname{I}[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $\operatorname{H}[S]$ describes the \key{complexity} of the communication system, and $\lambda \geq 0$ trades off the two aspects of efficiency.
While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\tree$.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
%The idea that efficient languages will permit the listener to infer grammati
Hence, substituting utterances $\utterance$ for signals $S$, and syntactic structures $\tree$ for underlying messages $R$, into (\ref{eq:eff-general}), we  arrive at the following efficiency metric for word order:
\begin{equation}\label{eq:efficiency-derived}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda \cdot R_\textit{Pred}
\end{equation}
where \key{parseability} is the amount of information that utterances provide about their underlying syntactic structures:
\begin{equation}
	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
and \key{predictability} is the negative entropy or surprisal of the language:
\begin{equation}
	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
\end{equation}
Parseability $\operatorname{I}[\utterance,\tree]$ is higher if utterances provide more information about their underlying syntactic structure.
Due to the identity $\operatorname{I}[\utterance, \tree] = \operatorname{H}[\tree] - \operatorname{H}[\tree|\utterance]$, parseability is maximized if every utterance can be parsed unambiguously -- that is, if the listener's uncertainty about syntactic structures given received utterances, $\operatorname{H}[\tree|\utterance]$, is zero.
Predictability $- \operatorname{H}[\utterance]$ is higher if the distribution over utterances is concentrated on a few utterances, and is maximized if there is just a single utterance.
It is also equal to the negative average of surprisal, which is a strong and linear predictor of human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.

@@

In the efficiency objective (\ref{eq:efficiency-derived})
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
\end{equation}
the value of $\lambda$ is constrained to be in $[0,1)$.
This means, surprisal must be weighted less strongly than parseability.

The reason is that greater values of $\lambda$ can mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
\operatorname{I}[\utterance; \tree] \leq \operatorname{H}[\utterance].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
%\begin{equation}
$R_{\textit{Eff}} = \operatorname{I}[\utterance; \tree] - \lambda \operatorname{H}[\utterance] \leq 0$,
%\end{equation}
and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $\operatorname{I}[\utterance; \tree]$ and $\operatorname{H}[\utterance]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars (see \emph{Materials and Methods}) precludes a collapse of all syntactic structures to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable description of the efficiency of communication systems.
In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.

@@

Greenberg's original correlations/implications

@@

Recall the definition of efficiency
\begin{equation}\label{eq:efficiency-recall}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
\end{equation}
where
\begin{equation}\label{eq:rpars}
	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
\end{equation}
\begin{equation}
	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
\end{equation}
where $t \sim \tree$ is the distribution over syntactic structures as found in databases of the language, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.

@@

	R_{Pars}^{\phi}(u,t) =  \log p_\phi(t|u)

@@

	R_{Pred}^{\psi}(u) = \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})

@@

	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi}

@@

R_{\textit{Eff}}^{\theta, \phi, \psi} := \E_{t \sim \mathcal{T}} \E_{u \sim p_{\mathcal{L}_\theta}(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]

@@

\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}

@@

Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures.

@@

Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.

@@

All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.

@@

\paragraph{Optimized Grammars}
As described in the main paper, for each language, we created 8 optimized languages for each optimization criterion.
We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
This maximizes statistical precision in detecting and quantifying correlations between the verb-object relation and other relations.

@@

Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers. These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.

@@

\subsection{n-gram language models}

We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.

Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
A bigram model predicts each word taking only the previous word into account.
This contrasts with LSTMs, which take the entire context into consideration.
Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.

@@

Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
We are addressing the following confound: It is conceivable that with enough data, our neural network language models and parsers would do equally well on real grammars and baseline grammars.
If the difference between random and real grammars is due to data sparsity in this way, then we expect that the difference will decrease as the amount of training data is increased.
If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference will persists as training data is increased.

We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
We considered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.

>>>>

We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding the objections by \citet{dunn2011evolved}, we refer to the follow-ups by \citet{levy2011computational}, and \citet{croft2011greenbergian}.}
Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
\citet{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
The verb--object relation corresponds to the  \emph{obj} relation defined by UD.
While most of the other relations also correspond to UD relations, some are not annotated reliably.
We were able formalize eleven out of Dryer's sixteen correlations in UD.
Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.

@@

\paragraph{Evaluating Accuracy of Formalization}
An anonymous reviewer notes that the mapping between Dryer's relations and UD is not perfect, since some of the UD relations subsume other relations.
Here we show that this is not a problem, since the ordering of the various relations subsumed under the UD label strongly agree typologically.

@@

We used the UD treebanks to confirm that different auxiliaries tend to pattern together, and that the most frequent order of the \emph{aux} relation coincides with that of inflected tense-aspect or negation auxiliaries.

We collected, for each UD language, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders, which we operationalized as their more common order in the treebank (auxiliary--head or head--auxiliary).
The dependency occurs in all but two very small treebanks (Telugu and Irish).
In 43 languages, all extracted auxiliaries had the same dominant order, with the possible exception of uninflected particles labeled \emph{aux} (Croatian, German, Polish, Ukrainian).
In three languages (Ancient Greek, Russian), there were other auxiliaries with different dominant order, but these were modal or passive auxiliaries.
Finally, in three languages (Afrikaans, Old Church Slavonic, and Persian), not all tense-aspect auxiliaries showed the same dominant order as the \emph{aux} dependency overall.
For instance, in Persian, the perfect auxiliary \emph{budan} follows the main verb, whereas the future auxiliary \emph{xaastan xaah-} precedes it~\citep[pp. 117, 121]{mace2015persian}.
%		(1) In Afrikaans, the past auxiliary \emph{het} shows dominant aux-head order, whereas the overall most common order of auxiliaries is head-aux (e.g. future auxiliary \emph{sal}).
%		Finally, in Old Church Slavonic, most aux dependencies have the reflexive \emph{sebe} which tends to precede the head, whereas inflected verbal auxiliaries (primarily, forms of \emph{byti}) have opposite order, leading to a seeming violation of the correlation  \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.

@@

\section{Formalizing Communicative Efficiency}
\subsection{Derivation and Relation to Other Work}

@@

Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms:
\begin{equation}\label{eq:eff-basic}
\text{Informativity} - \lambda \cdot \text{Complexity},
\end{equation}
with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely,hahn2018information, peloquin2019interactions,zaslavsky2019semantic}.

\paragraph{Derivation of our Formalization}
The basis for our precise formalization is the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
%The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms. %; there are some differences in the formalization of complexity.
If $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
\begin{equation}\label{eq:eff-general}
\operatorname{I}[S, R] - \lambda \cdot \operatorname{H}[S],
\end{equation}
where $\operatorname{I}[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $\operatorname{H}[S]$ describes the \key{complexity} of the communication system, and $\lambda \geq 0$ trades off the two aspects of efficiency.
While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\tree$.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
%The idea that efficient languages will permit the listener to infer grammati
Hence, substituting utterances $\utterance$ for signals $S$, and syntactic structures $\tree$ for underlying messages $R$, into (\ref{eq:eff-general}), we  arrive at the following efficiency metric for word order:
\begin{equation}\label{eq:efficiency-derived}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda \cdot R_\textit{Pred},
\end{equation}
where \key{parseability} is the amount of information that utterances provide about their underlying syntactic structures:
\begin{equation}
	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)},
\end{equation}
and \key{predictability} is the negative entropy or surprisal of the language:
\begin{equation}
	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u).
\end{equation}
Parseability $\operatorname{I}[\utterance,\tree]$ is higher if utterances provide more information about their underlying syntactic structure.
Due to the identity $\operatorname{I}[\utterance, \tree] = \operatorname{H}[\tree] - \operatorname{H}[\tree|\utterance]$, parseability is maximized if every utterance can be parsed unambiguously---that is, if the listener's uncertainty about syntactic structures given received utterances, $\operatorname{H}[\tree|\utterance]$, is zero.
Predictability $- \operatorname{H}[\utterance]$ is higher if the distribution over utterances is concentrated on a few utterances, and is maximized if there is just a single utterance.
It is also equal to the negative average of surprisal, which is a strong and linear predictor of human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.

@@

In the efficiency objective (\ref{eq:efficiency-derived})
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred},
\end{equation}
the value of $\lambda$ is constrained to be in $[0,1)$.
This means, surprisal must be weighted less strongly than parseability.

The reason is that greater values of $\lambda$ can mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
\operatorname{I}[\utterance; \tree] \leq \operatorname{H}[\utterance].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
%\begin{equation}
$R_{\textit{Eff}} = \operatorname{I}[\utterance; \tree] - \lambda \operatorname{H}[\utterance] \leq 0$,
%\end{equation}
and it takes the maximal possible value of zero if there is only a single utterance $\utterance$, in which case both $\operatorname{I}[\utterance; \tree]$ and $\operatorname{H}[\utterance]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars (see \emph{Materials and Methods}) precludes a collapse of all syntactic structures to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable description of the efficiency of communication systems.
In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.

@@

Greenberg's original correlations/implications % ?? TODO

@@

Recall the definition of efficiency
\begin{equation}\label{eq:efficiency-recall}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred},
\end{equation}
where
\begin{equation}\label{eq:rpars}
	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
\end{equation}
\begin{equation}
	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u),
\end{equation}
where $t \sim \tree$ is the distribution over syntactic structures as found in databases of the language, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.

@@

	R_{Pars}^{\phi}(u,t) =  \log p_\phi(t|u).

@@

	R_{Pred}^{\psi}(u) = \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1}).

@@

	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi},

@@

R_{\textit{Eff}}^{\theta, \phi, \psi} := \E_{t \sim \mathcal{T}} \E_{u \sim p_{\mathcal{L}_\theta}(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right].

@@

\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi},

@@

Note that the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures.

@@

Optimizing for only parseability (or predictability) is very similar---in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.

@@

All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
We do not use pretrained embeddings \citep{peters2018deep}; while these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.

@@

\paragraph{Optimized Grammars}
As described in the main paper, for each language, we created 8 optimized languages for each optimization criterion.
We enforced balanced distribution of object--verb and verb--object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
This maximizes statistical precision in detecting and quantifying correlations between the verb-object relation and other relations.

@@

Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers. These results show that our optimization process creates languages in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.

@@

\subsection{$n$-gram language models}

We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
In previous work, such as \cite{gildea2015human}, predictability has often been measured using $n$-gram models.

Here, we show that languages optimized for LSTM predictability are also optimal for $n$-gram predictability.
Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
A bigram model predicts each word taking only the previous word into account.
This contrasts with LSTMs, which take the entire context into consideration.
Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.

@@

Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
We are addressing the following confound: It is conceivable that with enough data, our neural network language models and parsers would do equally well on real grammars and baseline grammars.
If the difference between random and real grammars is due to data sparsity in this way, then we expect that the difference will decrease as the amount of training data is increased.
If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference will persist as training data is increased.

We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 million words), up to $\approx$ 300 times more data than is available for some other UD languages.
We considered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
